On Misuses of the Kolmogorov--Smirnov Test for One-Sample Goodness-of-Fit

### 2022-04-29 Manuscript/manuscript.tex

\begin{abstract}
The Kolmogorov--Smirnov (KS) test is one of the most popular goodness-of-fit tests 
for comparing a sample with a hypothesized parametric distribution. 
Nevertheless, it has often been misused. The standard one-sample KS test applies
to independent, continuous data with a hypothesized distribution that is 
completely specified. It is not uncommon, however, to see in the literature that
it was applied to dependent, discrete, or rounded data, with hypothesized 
distributions containing estimated parameters. For example, it has been 
``discovered" multiple times that the test is too conservative when the 
parameters are estimated. We demonstrate misuses of the one-sample KS test in 
three scenarios through simulation studies:
1) the hypothesized distribution has unspecified parameters;
2) the data are serially dependent; and
3) a combination of the first two scenarios.
For each scenario, we provide remedies for practical applications.
\end{abstract}

@@

The Kolmogorov-Smirnov (KS) test is one of the most popular goodness-of-fit 
tests for comparing a sample with a hypothesized parametric distribution.
Let $X_1, ..., X_n$ be a random sample from some continuous distribution. 
The null hypothesis $H_0$ is that the population distribution is $F$.
Let $F_n(\cdot)$ be the empirical cumulative distribution. The KS statistic is
\begin{equation}
  \label{eq:ks_standard}
  D = \sup_x | F_{n}(x) - F(x) |.
\end{equation}
The distribution of $D$ under $H_0$ turned out to be independent of the
distribution $F$ and a table of critical values for $D$ has been constructed
\citep{Massey} for various sample sizes $n$ and significance 
levels $\alpha$. If the value of $d$ exceeds the test's corresponding critical 
value, the null hypothesis is rejected. The KS test is available in popular 
statistical software packages, such as function \texttt{ks.test} in R 
\citep{R, Marsaglia}.
The standard one-sample KS test applies to independent data with a continuous
hypothesized distribution that is completely specified. In practice, however, it
has often been applied without realizing that one or more of these assumptions
do not hold.

@@

The standard KS test is not applicable when the hypothesized distribution 
contains fitted parameters. \citet{Steinskog} ``discovers'' the change in power 
when using estimated parameters and stress caution in using the KS test in such 
ways. \citet{Lilliefors} shows that using the standard table when values of the 
mean and standard deviation are estimated obtains extremely conservative 
results. This is supported by \citet{Capasso}, who concludes that failing to 
re-estimate the parameters may lead to wrong, overly-conservative approximations 
to the distributions of goodness-of-fit test statistics based on the empirical 
distribution function. \citet{Capasso} also notes that the impact of this 
possible mistake may turn out to be dramatic and does not vanish as the sample 
size increases. Remedies are provided by \citet{Babu} and \citet{Genest} in the
form of bootstrap.  \citet{Babu} details the bootstrap
procedure for goodness-of-fit tests and notes that both parametric and 
non-parametric procedures lead to correct asymptotic levels, however there 
is a correction required for the non-parametric case.
\citet{Genest} provides validity for using parametric 
bootstrap with various goodness-of-fit tests.

@@

Nonparametric bootstrap can also be used to approximate the null distribution 
of the testing statistic. The procedure is similar, however the resampling is 
performed with the empirical distribution instead of the fitted parametric
distribution and there is a correction for bias that is required
\citep{Babu}.
\begin{enumerate}
  \item 
    Draw a random sample $X_1^*,...,X_n^*$ from the empirical distribution $F_n$
    with replacement
  \item 
    Fit $F_\theta$ to the sample and obtain estimated $\hat\theta_n^*$
  \item
    Obtain the empirical distribution function of the random sample $F_n^*$
  \item 
    Calculate bootstrap KS statistic
    \[
      D^* = \sup_x | F_n^* (x)- F_{\hat\theta_n}^*(x) - B_n(x) |.
    \]
    where $B_{n}(x) = \sqrt{n}(F_{n}(x) - F_{\hat\theta_n}(x))$ is the known 
    bias term \citep{Babu}
  \item
    Repeat the previous steps a large number $B$ times and use the empirical 
    distribution of $D^*$ to approximate the null distribution of the observed 
    statistic.
\end{enumerate}
The p-value can be calculated by counting the number of bootstrap KS 
statistics greater than or equal to the observed KS statistic, and then dividing 
by the number of bootstrap samples. Figure~\ref{fig:hist_fitted} displays the 
results of from our simulations. We would expect the distribution of p-values 
to be normal in the case where the KS test holds its size. It is clear from the 
figure that both parametric and nonparametric bootstrap processes correct for 
the problem of fitted parameters. The plots for the bootstrapped p-values appear
to be $U(0,1)$, unlike the naive p-values.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_fitted}
  \caption{The Naive plot (left) is the histogram of p-values from the standard 
  KS test with fitted parameters. The Parametric plot (middle) is the histogram
  of p-values from implementing parametric bootstrap. The Babu plot (right) is 
  the histogram of p-values from implementing the nonparametric bootstrap with a 
  correction for bias. In each case, $1000$ replicate tests were performed on 
  the standard normal distribution with sample size $n = 100$. $B = 1000$ 
  bootstrap samples are obtained for each test.}
  \label{fig:hist_fitted}
\end{figure}

@@

The KS test also displays issues in the case of dependent data. As mentioned, an 
assumption of the test is that the data is independent. Unfortunately, real data
is often temporally or spatially dependent and the results of a goodness-of-fit 
test would be valuable. When the KS test is performed on dependent data it 
performs poorly. In this situation, the null hypothesis is $H_0$: the random 
sample $X_1, \ldots, X_n$ come from a continuous distribution $F$ where $F$ has 
completely specified parameters. The test statistic is the same as 
Equation~\ref{eq:ks_standard}. However, since the data is dependent, the null 
hypothesis is wrong and must be corrected. This is demonstrated with a 
simulation in Figure~\ref{fig:hist_correlation}. Data is generated from a 
first-order autoregressive model (AR(1)) with a standard normal distribution. 
P-values are gathered from $1000$ replicate tests for different levels of the 
AR coefficient $\psi$, varying from $(-3,3)$. In the presence of serially 
dependent data, the distribution of p-values no longer follows $U(0, 1)$ as 
would be expected of a valid test. The results echo those of \citet{Durilleul} 
that the KS statistic is too liberal for positive autocorrelation values, and 
that the behavior is asymmetrical for negative values.

\begin{figure}[tbp]
  \centering
  \includegraphics{hist_correlation}
  \caption{Each histogram is a plot of the p-values of standard KS tests
  performed on dependent data simulated from an AR(1) process of the standard 
  normal distribution. The titles represent the various levels of the AR 
  coefficient $\psi$ tested. The sample size is $n = 100$ and $1000$ replicate 
  tests were performed for each AR coefficient.}
  \label{fig:hist_correlation}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics{hist_ar1_D}
  \caption{The Naive plot (left) is the histogram of p-values from the 
  standard KS test. The Copula plot (right) is the histogram of p-values from 
  implementing parametric bootstrap with copulas for dependence. In each case, 
  $1000$ replicate tests were performed with the data generated from an AR(1) 
  process of the standard normal distribution with $\psi = .5$. The sample size 
  is $n = 100$ and $B = 1000$ bootstrap samples are obtained for each test.}
  \label{fig:hist_ar1_D}
\end{figure}

Figure~\ref{fig:hist_ar1_D} displays the results of the copula remedy for 
dependent data. The data is generated from the standard normal distribution with
an AR(1) dependence structure where $\psi = .5$. The copula used is to model 
dependence is the normal copula. The distribution of p-values is again expected
to be $U(0, 1)$ for a valid test. The Naive plot is clearly not uniform and 
reinforces the results shown in Figure~\ref{fig:hist_correlation}. The Copula 
plot, which implements the aforementioned procedure to correct for dependence
in the data, appears to be uniform and restores the size of the KS test.
\jy{we didn't study the power here (probability to reject the null when the null
  is false). Can do it for the paper.}

\begin{figure}[tbp]
  \centering
  \includegraphics{hist_ma1_arma_ar2_D}
  \caption{The Naive plots are the histograms of p-values from the standard KS 
  test. The Copula plots are the histograms of p-values from parametric 
  bootstrap with copulas for dependence. The data is generated from an MA(1) 
  process (left) with $\theta = .5$, ARMA(1, 1) process (middle) with $\psi = 
  .5$ and $\theta = .3$, and AR(2) process (right) with $\psi = (.5, .3)$ of the 
  standard normal distribution. In each case, $1000$ replicate tests were 
  performed with sample size $n = 100$ and $B = 1000$ bootstrap samples.}
  \label{fig:hist_ma1_arma_ar2_D}
\end{figure}

@@

\jy{I like the facet grid version better}

\begin{figure}[tbp]
  \centering
  \includegraphics{hist_gamma_gev_D}
  \caption{The Naive plots are the histograms of p-values from the standard KS 
  test. The Copula plots are the histograms of p-values from parametric 
  bootstrap with copulas for dependence. The data is generated from $Gamma(3, 
  1)$ (left) and $GEV(0, .2, 1)$ (right) with a correlation coefficient of $r = 
  0.5$. In each case, $1000$ replicate tests were performed with sample size $n 
  = 100$ and $B = 1000$ bootstrap samples.}
  \label{fig:hist_gamma_gev_D}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics{hist_ar1_FD}
  \caption{The Super Naive plot (left) is the histogram of p-values from the 
  standard KS test with fitted parameters. The Naive Parametric plot (middle) is
  the histogram of p-values from implementing parametric bootstrap. The Copula 
  plot is the histogram of p-values from implementing parametric bootstrap with 
  copulas for dependence and correcting for fitted parameters. In each case, 
  $1000$ replicate tests were performed with the data generated from an AR(1) 
  process of the standard normal distribution with $\psi = .5$. The sample size 
  is $n = 100$ and $B = 1000$ bootstrap samples are obtained for each test.}
  \label{fig:hist_ar1_FD}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics{hist_ma1_arma_ar2_FD}
  \caption{The Super Naive plots (left) are the histograms of p-values from the 
  standard KS test with fitted parameters. The Naive Parametric plots (middle) 
  are the histograms of p-values from implementing parametric bootstrap. The 
  Copula plots are the histograms of p-values from implementing parametric 
  bootstrap with copulas for dependence and correcting for fitted parameters. 
  The data is generated from an MA(1) process (left) with $\theta = .5$, ARMA(1,
  1) process (middle) with $\psi = .5$ and $\theta = .3$, and AR(2) process 
  (right) with $\psi = (.5, .3)$ of the standard normal distribution. In each 
  case, $1000$ replicate tests were performed with sample size $n = 100$ and 
  $B = 1000$ bootstrap samples.}
  \label{fig:hist_ma1_arma_ar2_FD}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics{hist_gamma_gev_FD}
  \caption{The Super Naive plots (left) are the histograms of p-values from the 
  standard KS test with fitted parameters. The Naive Parametric plots (middle) 
  are the histograms of p-values from implementing parametric bootstrap. The 
  Copula plots are the histograms of p-values from implementing parametric 
  bootstrap with copulas for dependence and correcting for fitted parameters. 
  The data is generated from $Gamma(3, 1)$ (left) and $GEV(0, .2, 1)$ (right) 
  with a correlation coefficient of $r = 0.5$. In each case, $1000$ replicate 
  tests were performed with sample size $n = 100$ and $B = 1000$ bootstrap 
  samples.}
  \label{fig:hist_gamma_gev_FD}
\end{figure}

>>>>

\begin{abstract}
The Kolmogorov--Smirnov (KS) test is one of the most popular goodness-of-fit tests 
for comparing a sample with a hypothesized parametric distribution. 
Nevertheless, it has often been misused. The standard one-sample KS test applies
to independent, continuous data with a hypothesized distribution that is 
completely specified. It is not uncommon, however, to see in the literature that
it was applied to dependent, discrete, or rounded data, with hypothesized 
distributions containing estimated parameters. For example, it has been 
``discovered'' multiple times that the test is too conservative when the 
parameters are estimated. We demonstrate misuses of the one-sample KS test in 
three scenarios through simulation studies:
1) the hypothesized distribution has unspecified parameters;
2) the data are serially dependent; and
3) a combination of the first two scenarios.
For each scenario, we provide remedies for practical applications.
\end{abstract}

@@

The Kolmogorov-Smirnov (KS) test is one of the most popular goodness-of-fit 
tests for comparing a sample with a hypothesized parametric distribution.
Let $X_1, ..., X_n$ be a random sample from some continuous distribution. 
The null hypothesis $H_0$ is that the population distribution is $F$.
Let $F_n(\cdot)$ be the empirical cumulative distribution. The KS statistic is
\begin{equation}
  \label{eq:ks_standard}
  D = \sup_x | F_{n}(x) - F(x) |.
\end{equation}
The distribution of $D$ under $H_0$ turns out to be independent of the
distribution $F$ and a table of critical values for $D$ has been constructed
\citep{Massey} for various sample sizes $n$ and significance 
levels $\alpha$. If the value of $D$ exceeds the test's corresponding critical 
value the null hypothesis is rejected. The KS test is available in popular 
statistical software packages, such as function \texttt{ks.test} in R 
\citep{R, Marsaglia}.

The standard one-sample KS test applies to independent data with a continuous
hypothesized distribution that is completely specified. In practice, however, it
has often been applied without realizing that one or more of these assumptions
do not hold.

@@

The standard KS test is not applicable when the hypothesized distribution 
contains fitted parameters. \citet{Steinskog} ``discovers'' the change in power 
when using estimated parameters and stresses caution in using the KS test in 
such ways. \citet{Lilliefors} shows for the normal distribution that using the 
standard table when values of the mean and standard deviation are estimated 
obtains extremely conservative results. This is supported by \citet{Capasso}, 
who concludes that failing to re-estimate the parameters may lead to wrong, 
overly-conservative approximations to the distributions of goodness-of-fit test 
statistics based on the empirical distribution function. \citet{Capasso} also 
notes that the impact of this mistake may turn out to be dramatic and does not 
vanish as the sample size increases. Remedies are provided by \citet{Babu} and 
\citet{Genest} in the form of bootstrap. \citet{Babu} details the bootstrap
procedure for goodness-of-fit tests and notes that both parametric and 
non-parametric procedures lead to correct asymptotic levels, however there is a 
correction required for the non-parametric case. \citet{Genest} provides 
validity for using parametric bootstrap with various goodness-of-fit tests.

@@

Nonparametric bootstrap can also be used to approximate the null distribution 
of the testing statistic. The procedure is similar, however the resampling is 
performed with the empirical distribution instead of the fitted parametric
distribution and there is a correction for bias that is required
\citep{Babu}.
\begin{enumerate}
  \item 
    Draw a random sample $X_1^*,...,X_n^*$ from the empirical distribution $F_n$
    with replacement
  \item 
    Fit $F_\theta$ to the sample and obtain estimated $\hat\theta_n^*$
  \item
    Obtain the empirical distribution function of the random sample $F_n^*$
  \item 
    Calculate bootstrap KS statistic
    \[
      D^* = \sup_x | F_n^* (x)- F_{\hat\theta_n}^*(x) - B_n(x) |.
    \]
    where $B_{n}(x) = \sqrt{n}(F_{n}(x) - F_{\hat\theta_n}(x))$ is the known 
    bias term \citep{Babu}
  \item
    Repeat the previous steps a large number $B$ times and use the empirical 
    distribution of $D^*$ to approximate the null distribution of the observed 
    statistic.
\end{enumerate}
The p-value can be calculated by counting the number of bootstrap KS 
statistics greater than or equal to the observed KS statistic, and then dividing 
by the number of bootstrap samples. Figure~\ref{fig:hist_fitted} displays the 
results of from our simulations. We would expect the distribution of p-values 
to be uniform in the case where the KS test holds its size. It is clear from the 
figure that both parametric and nonparametric bootstrap processes correct for 
the problem of fitted parameters. The plots for the bootstrapped p-values appear
to be $U(0,1)$, unlike the naive p-values.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_fitted}
  \caption{The Naive plot (left) is the histogram of p-values from the standard 
  KS test with fitted parameters. The Parametric plot (middle) is the histogram
  of p-values from implementing parametric bootstrap. The Babu plot (right) is 
  the histogram of p-values from implementing nonparametric bootstrap with a 
  correction for bias. In each case, $1000$ replicate tests were performed on 
  the standard normal distribution with sample size $n = 100$. $B = 1000$ 
  bootstrap samples are obtained for each test.}
  \label{fig:hist_fitted}
\end{figure}

@@

The KS test also displays issues in the case of dependent data. As mentioned, an 
assumption of the test is that the data is independent. Unfortunately, real data
is often temporally or spatially dependent and the results of a goodness-of-fit 
test would be valuable. When the KS test is performed on dependent data it 
performs poorly. In this situation, the null hypothesis is $H_0$: the random 
sample $X_1, \ldots, X_n$ come from a continuous distribution $F$ where $F$ has 
completely specified parameters. The test statistic is the same as 
Equation~\eqref{eq:ks_standard}. However, since the data is dependent, the null 
hypothesis is wrong and must be corrected. This is demonstrated with a 
simulation in Figure~\ref{fig:hist_correlation}. Data is generated from a 
first-order autoregressive model (AR(1)) with a standard normal distribution. 
P-values are gathered from $1000$ replicate tests for different levels of the 
AR coefficient $\psi$, varying from $(-3,3)$. In the presence of serially 
dependent data, the distribution of p-values no longer follows $U(0, 1)$ as 
would be expected of a valid test. The results echo those of \citet{Durilleul} 
that the KS statistic is too liberal for positive autocorrelation values, and 
that the behavior is asymmetrical for negative values.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_correlation}
  \caption{Each histogram is a plot of the p-values of standard KS tests
  performed on dependent data simulated from an AR(1) process of the standard 
  normal distribution. The titles represent the various levels of the AR 
  coefficient $\psi$ tested. The sample size is $n = 100$ and $1000$ replicate 
  tests were performed for each AR coefficient.}
  \label{fig:hist_correlation}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_ar1_D}
  \caption{The Naive plot (left) is the histogram of p-values from the 
  standard KS test. The Copula plot (right) is the histogram of p-values from 
  implementing parametric bootstrap with copulas for dependence. In each case, 
  $1000$ replicate tests were performed with the data generated from an AR(1) 
  process of the standard normal distribution with $\psi = .5$. The sample size 
  is $n = 100$ and $B = 1000$ bootstrap samples are obtained for each test.}
  \label{fig:hist_ar1_D}
\end{figure}

Figure~\ref{fig:hist_ar1_D} displays the results of the copula remedy for 
dependent data. The data is generated from the standard normal distribution with
an AR(1) dependence structure where $\psi = .5$. The copula used is to model 
dependence is the normal copula. The distribution of p-values is again expected
to be $U(0, 1)$ for a valid test. The Naive plot is clearly not uniform and 
reinforces the results shown in Figure~\ref{fig:hist_correlation}. The Copula 
plot, which implements the aforementioned procedure to correct for dependence
in the data, appears to be uniform and restores the size of the KS test.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_ma1_arma_ar2_D}
  \caption{The Naive plots are the histograms of p-values from the standard KS 
  test. The Copula plots are the histograms of p-values from parametric 
  bootstrap with copulas for dependence. The data is generated from an MA(1) 
  process (left) with $\theta = .5$, ARMA(1, 1) process (middle) with $\psi = 
  .5$ and $\theta = .3$, and AR(2) process (right) with $\psi = (.5, .3)$ of the 
  standard normal distribution. In each case, $1000$ replicate tests were 
  performed with sample size $n = 100$ and $B = 1000$ bootstrap samples.}
  \label{fig:hist_ma1_arma_ar2_D}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_gamma_gev_D}
  \caption{The Naive plots are the histograms of p-values from the standard KS 
  test. The Copula plots are the histograms of p-values from parametric 
  bootstrap with copulas for dependence. The data is generated from $Gamma(3, 
  1)$ (left) and $GEV(0, .2, 1)$ (right) with a correlation coefficient of $r = 
  0.5$. In each case, $1000$ replicate tests were performed with sample size $n 
  = 100$ and $B = 1000$ bootstrap samples.}
  \label{fig:hist_gamma_gev_D}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_ar1_FD}
  \caption{The Super Naive plot (left) is the histogram of p-values from the 
  standard KS test with fitted parameters. The Naive Parametric plot (middle) is
  the histogram of p-values from implementing parametric bootstrap. The Copula 
  plot is the histogram of p-values from implementing parametric bootstrap with 
  copulas for dependence and correcting for fitted parameters. In each case, 
  $1000$ replicate tests were performed with the data generated from an AR(1) 
  process of the standard normal distribution with $\psi = .5$. The sample size 
  is $n = 100$ and $B = 1000$ bootstrap samples are obtained for each test.}
  \label{fig:hist_ar1_FD}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_ma1_arma_ar2_FD}
  \caption{The Super Naive plots (left) are the histograms of p-values from the 
  standard KS test with fitted parameters. The Naive Parametric plots (middle) 
  are the histograms of p-values from implementing parametric bootstrap. The 
  Copula plots (bottom) are the histograms of p-values from implementing 
  parametric bootstrap with copulas for dependence and correcting for fitted 
  parameters. The data is generated from an MA(1) process (left) with 
  $\theta = .5$, ARMA(1, 1) process (middle) with $\psi = .5$ and $\theta = .3$,
  and AR(2) process (right) with $\psi = (.5, .3)$ of the standard normal 
  distribution. In each case, $1000$ replicate tests were performed with sample 
  size $n = 100$ and $B = 1000$ bootstrap samples.}
  \label{fig:hist_ma1_arma_ar2_FD}
\end{figure}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{hist_gamma_gev_FD}
  \caption{The Super Naive plots (left) are the histograms of p-values from the 
  standard KS test with fitted parameters. The Naive Parametric plots (middle) 
  are the histograms of p-values from implementing parametric bootstrap. The 
  Copula plots are the histograms of p-values from implementing parametric 
  bootstrap with copulas for dependence and correcting for fitted parameters. 
  The data is generated from $Gamma(3, 1)$ (left) and $GEV(0, .2, 1)$ (right) 
  with a correlation coefficient of $r = 0.5$. In each case, $1000$ replicate 
  tests were performed with sample size $n = 100$ and $B = 1000$ bootstrap 
  samples.}
  \label{fig:hist_gamma_gev_FD}
\end{figure}
